---
title: "mauer"
output: html_document
date: "2024-07-04"
---

# load libraries

```{r}

if (!requireNamespace("sf", quietly = TRUE)) install.packages("sf")
if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")

library(labelr) 
library(tidyverse)
library(janitor)
library(lubridate)
library(ggplot2)
library(readxl)
library(writexl)
library(dplyr)
library(sf)


```

# load data

```{r}
maueratbat <- read_csv("maueratbat.csv")

```

# remove the extra plate appearance + add a column for home and away + fip the y-axis so charts are right side up
# row with 2007-09-04 00:00:00 in the gamedate column

```{r}

maueratbatclean <- maueratbat %>%
  clean_names() %>%
  # remove the extra plate appearance
  filter(!grepl("2007-09-04", game_date)) %>%
  # add a column for home and away
  mutate(location = ifelse(grepl("Twins$", game), "home", "away")) %>%
  # flip the y-axis
  mutate(ball_in_play_y_flipped = -ball_in_play_y) %>%
  # make dates actual date values
  mutate(game_date = as.Date(game_date)) %>%
  mutate(inning = gsub("[^0-9]", "", inning))

```

# only show me rows where the ball was in play

```{r}

mauerinplay <- maueratbatclean %>%
  filter(!(ball_in_play_x == 0 & ball_in_play_y == 0))

```

# We probably should think about creating a new field that simplifies the “result” field into something with fewer categories

```{r}

mauerresultcount <- maueratbatclean %>%
  group_by(result) %>%
  count()

mauersimplified <- maueratbatclean %>%
mutate(simple_result = case_when(
    result %in% c('Strikeout Looking', 'Strikeout Swinging', 'Dropped Third Strike', 'Dropped Third Strike (PB)', 'Dropped Third Strike (WP)') ~ 'strikeout',
    result %in% c('Home Run', 'Inside the Park HR') ~ 'home run',
    result %in% c('Walked', 'Intentionally Walked') ~ 'walked',
    result %in% c('Sacrifice', 'Sacrifice Fly', 'Sacrifice Fly Error') ~ 'sacrifice',
    result %in% c('Double', 'Ground Rule Double') ~ 'double',
    result %in% c('Grounded Into Double Play', 'Non-Force GDP', 'Double Play') ~ 'turn two',
    result %in% c('Fly Out', 'Ground Out', 'Line Drive') ~ 'out',
    result %in% c('Foul Fly Out') ~ 'foul',
    result %in% c("Fielder's Choice", "Fielder's Choice+Error") ~ 'fielders choice',
    result %in% c('Triple') ~ 'triple',
    result %in% c('Single') ~ 'single',
    result %in% c('Error', 'Assist With Error') ~ 'outfield error',
    result %in% c('Hit By Pitch') ~ 'hit by pitch',
    TRUE ~ result  # This handles any other values that don't match the above conditions
  ))


```

# transform the coordinates for 2004-05 data

```{r}

# Create dataframe of just 2004-2005 rows
mauer2004_05 <- mauersimplified %>%
  filter(year(game_date) %in% c(2004, 2005))

write_xlsx(mauer2004_05, path = "mauer2004_05.xlsx")

# Function to transform coordinates from old system to new system
transform_coordinates <- function(x_old, y_old) {
  # 2004-2005 system
  feet_per_pixel_2004 = 3.6365
  home_plate_x_2004 = 82.5
  home_plate_y_2004 = 120.5
  
  # Convert old pixels to feet
  x_feet = (x_old - home_plate_x_2004) * feet_per_pixel_2004
  y_feet = (y_old - home_plate_y_2004) * feet_per_pixel_2004
  
  # 2006-present system
  feet_per_pixel_2006 = 1.1068
  home_plate_x_2006 = 288
  home_plate_y_2006 = 405
  
  # Convert feet to new pixels
  x_new = x_feet / feet_per_pixel_2006 + home_plate_x_2006
  y_new = y_feet / feet_per_pixel_2006 + home_plate_y_2006
  
  return(data.frame(x_new = x_new, y_new = y_new))
}

# Adjust coordinates for 2004-2005 data
newmauer2004_05 <- mauer2004_05 %>%
  rowwise() %>%
  mutate(
    coords = list(transform_coordinates(ball_in_play_x, ball_in_play_y_flipped))
  ) %>%
  unnest(cols = c(coords))

```

# now I am just making a 2006-present dataframe to join with the new 2004-05 dataframe

```{r}

mauer2006_present <- mauersimplified %>%
  filter(!(year(game_date) %in% c(2004, 2005))) %>%
  mutate(x_new = ball_in_play_x) %>%
  mutate(y_new = ball_in_play_y_flipped)

```

# now I am joining the two dataframes

```{r}

adjustedpixels <- rbind(newmauer2004_05, mauer2006_present)

#only balls in play for adjusted pixels
adjustedpixels <- adjustedpixels %>%
filter(!(ball_in_play_x == 0 & ball_in_play_y == 0))


# write this to excel
write_xlsx(adjustedpixels, path = "maueratbat_pixels.xlsx")

```

# did mauer improve on getting balls in play?
# not really lol

```{r}

# Calculate counts by season and condition
counts <- mauersimplified %>%
  mutate(
    condition = case_when(
      ball_in_play_x == 0 & ball_in_play_y_flipped == 0 ~ "Both Zero",
      TRUE ~ "Either Non-zero"
    )
  ) %>%
  group_by(season, condition) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  pivot_wider(names_from = condition, values_from = count, values_fill = 0)


```

# home vs away aka does he love minnesota sooo much
# no 

```{r}

homevaway <- mauersimplified %>%
  filter(grepl("home run", simple_result)) %>%
  group_by(season, location, simple_result) %>%
  count()


```

# homeruns over time 

```{r}

homerunbyseason <- mauersimplified %>%
  #filter(season == "2009") %>%
  filter(simple_result == "home run") %>%
  group_by(season) %>%
  count()

ggplot(homerunbyseason, aes(x = season, y = n)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Home Runs by Season",
    x = "Season",
    y = "Count of Home Runs"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```
# triples over time

```{r}

triplesbyseason <- mauersimplified %>%
  #filter(season == "2009") %>%
  filter(simple_result == "triple") %>%
  group_by(season) %>%
  count()

ggplot(triplesbyseason, aes(x = season, y = n)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Triples by Season",
    x = "Season",
    y = "Count of Triples"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```
# doubles by season

```{r}

doublesbyseason <- mauersimplified %>%
  #filter(season == "2009") %>%
  filter(simple_result == "double") %>%
  group_by(season) %>%
  count()

ggplot(doublesbyseason, aes(x = season, y = n)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Doubles by Season",
    x = "Season",
    y = "Count of Doubles"
  ) +
  theme_minimal() +
  theme(legend.position = "none")


```
# singles over time

```{r}

singlesbyseason <- mauersimplified %>%
  #filter(season == "2009") %>%
  filter(simple_result == "triple") %>%
  group_by(season) %>%
  count()

ggplot(singlesbyseason, aes(x = season, y = n)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Singles by Season",
    x = "Season",
    y = "Count of Singles"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```
# outs over time

```{r}

outsbyseason <- mauersimplified %>%
  #filter(season == "2009") %>%
  filter(simple_result == "out") %>%
  group_by(season) %>%
  count()

ggplot(outsbyseason, aes(x = season, y = n)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Outs by Season",
    x = "Season",
    y = "Count of Outs"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

# fouls by season

```{r}

foulsbyseason <- mauersimplified %>%
  #filter(season == "2009") %>%
  filter(simple_result == "foul") %>%
  group_by(season) %>%
  count()

ggplot(foulsbyseason, aes(x = season, y = n)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Fouls by Season",
    x = "Season",
    y = "Count of Fouls"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```


# strikeouts by season

```{r}

strikeoutsbyseason <- mauersimplified %>%
  #filter(season == "2009") %>%
  filter(simple_result == "strikeout") %>%
  group_by(season) %>%
  count()

ggplot(strikeoutsbyseason, aes(x = season, y = n)) +
  geom_bar(stat = "identity") +
  labs(
    title = "strikeouts by Season",
    x = "Season",
    y = "Count of strikeouts"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```
# The unusualness of the 2009 season - his best, MVP, started late

```{r}
homerun2009 <- mauersimplified %>%
  filter(simple_result == "home run") %>%
  group_by(season, simple_result) %>%
  count()

all2009 <- mauersimplified %>%
  filter(season == "2009")

write_xlsx(homerun2009, path = "homerun2009.xlsx")


```

# womp womp gavin floyd

```{r}

sorrygavin <- mauersimplified %>%
  filter(pitcher == "Gavin Floyd") %>%
  #filter(simple_result == "home run") %>%
  group_by(simple_result) %>%
  count()

verlander <- mauersimplified %>%
  filter(pitcher == "Justin Verlander")
  #group_by(simple_result) %>%
  #count()

strikeouts <- mauersimplified %>%
  filter(simple_result == "strikeout") %>%
  group_by(pitcher, season) %>%
  count()

```

# nice job Sabathia

```{r}

sabathia <- mauersimplified %>%
  filter(pitcher == "CC Sabathia") %>%
  group_by(game_date)

```


# When would a hit be categorized as a line drive rather than a single or double or out or something? Does that just indicate he hit it but didn't score on it?

# Can we make sure I accurately grouped things? Any line that I commented on we should go over
